---
title: "List of methods"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
---

## Highest averages methods

|Method|Argument|Rule|Approx. first values|
|:---|:---|:---|:---|
|D'hondt / Jefferson |`"d'hondt"`, `"dt"`, or `"jefferson"`|$r = k + 1$|1, 2, 3, 4, ...|
|Adams / Cambridge|`"adams"` or `"cambridge"`|$r = k$|0, 1, 2, 3, ...|
|Sainte-Laguë / Webster|`"sl"`, `"sainte-lague"`, or `"webster"`|$r = k + 0.5$|0.5, 1.5, 2.5, 3.5, ...|
|Modified Sainte-Laguë|`"msl"`|$r = k + 0.5, r_1 = 0.7$|0.7, 1.5, 2.5, 3.5, ...|
|Danish|`"danish"`|$r = k + \frac{1}{3}$|0.33, 1.33, 2.33, 3.33, ...|
|Imperiali|`"imperiali"`|$r = k + 2$|2, 3, 4, 5, ...|
|Huntington-Hill|`"hh"` or `"huntington-hill"`|$r = \sqrt{k (k + 1)}$|0, 1.41, 2.45, 3.46, ...|
|Dean|`"dean"`|$\frac{2}{\frac{1}{k} + \frac{1}{k + 1}}$||
|$\alpha$-divergence|`"ad"`|See @nte-ad |See @nte-ad |
|Custom|`"custom"`|See @nte-custom |See @nte-custom |

:::{#nte-ad .callout-note collapse="true"}
## About $\alpha$-divergence

Divisor ($r$) is calculated as follow:

$$
r = \begin{cases}
\frac{1}{e} \frac{(k + 1)^{k + 1}}{k^k}& \mbox{if}\quad\alpha = 1,\\
\frac{1}{\log{\frac{k + 1}{k}}} & \mbox{if}\quad\alpha = 0,\\
\Bigl(\frac{(k + 1)^\alpha - k^\alpha}{\alpha}\Bigr)^{\frac{1}{\alpha - 1}} & \mbox{otherwise.}\\
\end{cases}
$$

|$\alpha$|Method|
|---:|:---|
|+$\infty$|D'hondt / Jefferson|
|3|$\frac{1}{3}$|
|2|Sainte-Laguë / Webster|
|1|Theil|
|0|Theil and Schrage|
|-1|$\simeq$Hill|
|-4|Dean|
|-$\infty$|Adams / Cambridge|

Please note that extra parameter $\alpha$ is required (`extra`). 

:::{.panel-tabset}
## Example 1

```{r}
#| message: false
library(PRcalc)
data(jp_upper_2019)
# Jefferson vs alpha = 10
list(prcalc(jp_upper_2019, m = 50, method = "dt"),
     prcalc(jp_upper_2019, m = 50, method = "ad", extra = 10)) |> 
  compare() |> 
  print(use_gt = TRUE)
```

## Example 2

```{r}
#| message: false
# Dean vs alpha = -4
list(prcalc(jp_upper_2019, m = 50, method = "dean"),
     prcalc(jp_upper_2019, m = 50, method = "ad", extra = -4)) |> 
  compare() |> 
  print(use_gt = TRUE)
```
:::

:::

:::{#nte-custom .callout-note collapse="true"}
## About `"custom"` method

Please note that extra parameter $\alpha$ is required (`extra`). `extra` must be length of maximum value of `m`.

```{r}
#| message: false
list(prcalc(jp_upper_2019, m = 50, method = "msl"),
     prcalc(jp_upper_2019, m = 50, method = "custom", 
            extra = c(1.4, seq(3, 100, by = 2)))) |> 
  compare() |> 
  print(use_gt = TRUE)
```
:::

## Largest reminder method

|Method|Argument|Quota|
|:---|:---|:---|:---|
|Hare-Niemeyer|`"hn"`, `"hare"`, or `"hare-niemeyer"`||
