---
title: "Comparison with other packages"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
    collapse: true
---

## Setting

```{r}
#| message: false
library(tidyverse)
library(PRcalc)
library(seatdist)

dp <- disproportionality

obj1 <- prcalc(jp_upper_2019, m = 50, method = "dt")
ind1 <- index(obj1)
v1   <- obj1$raw$Vote
s1   <- obj1$dist$Vote

obj2 <- prcalc(us_census_2020, m = 435, method = "hh")
ind2 <- index(obj2)
v2   <- obj2$raw$Pop
s2   <- obj2$dist$Pop
```

## Comparison

### D'Hondt

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["dhondt"]
ind2["dhondt"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "dhondt")$value
dp(s = s2, v = v2, measure = "dhondt")$value
```
:::
::::

### Monroe

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["monroe"]
ind2["monroe"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "monroe")$value
dp(s = s2, v = v2, measure = "monroe")$value
```
:::
::::

### Maximum absolute deviation

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["maxdev"]
ind2["maxdev"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "monroe")$value
dp(s = s2, v = v2, measure = "monroe")$value
```
:::
::::

### Max-Min ratio

```{r}
#| filename: "{PRcalc}"
ind1["mm_ratio"]
ind2["mm_ratio"]
```


### Rae

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["maxdev"]
ind2["maxdev"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "maxdev")$value
dp(s = s2, v = v2, measure = "maxdev")$value
```
:::
::::

### Loosemore & Hanby

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["lh"]
ind2["lh"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "loosemore hanby")$value
dp(s = s2, v = v2, measure = "loosemore hanby")$value
```
:::
::::

### Grofman

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["grofman"]
ind2["grofman"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "grofman")$value
dp(s = s2, v = v2, measure = "grofman")$value
```
:::
::::

### Lijphart

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["lijphart"]
ind2["lijphart"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "lijphart")$value
dp(s = s2, v = v2, measure = "lijphart")$value
```
:::
::::

Since the results for each package are different, we need to be sure we know which results are correct. Lijphart index is calculated as follow:

$$
\mbox{I}_{\mbox{Lijphart}} = \frac{|s_a - v_a| + |s_b - v_b|}{2},
$$

where $s_a$ and $v_a$ are the proportion of seats and votes for the party that received the largest number of votes, and $s_b$ and $v_b$ are that for the party that received the second largest of votes. We can compute the above equation directly.

```{r}
new_s1 <- s1[order(v1, decreasing = TRUE)]
new_v1 <- v1[order(v1, decreasing = TRUE)]

new_s1 <- new_s1 / sum(new_s1)
new_v1 <- new_v1 / sum(new_v1)

new_s1
new_v1

(abs(new_s1[1] - new_v1[1]) + abs(new_s1[2] - new_v1[2])) / 2
```

The calculation results show that {PRcalc} is correct.

### Gallagher

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["gallagher"]
ind2["gallagher"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "gallagher")$value
dp(s = s2, v = v2, measure = "gallagher")$value
```
:::
::::

### Generalized Gallagher

* Parameter $k$ (`k`) is required. Default is 2.

::::{.columns}
:::{.column width=49%}
```{r}
#| filename: "{PRcalc}"
index(obj1, k = 2)["g_gallagher"]
index(obj2, k = 3)["g_gallagher"]
```
:::

:::{.column width=2%}
:::

:::{.column width=49%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, k = 2, 
   measure = "kindex")$value
dp(s = s2, v = v2, k = 3, 
   measure = "kindex")$value
```
:::
::::

### Gatev

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["gatev"]
ind2["gatev"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "gatev")$value
dp(s = s2, v = v2, measure = "gatev")$value
```
:::
::::

### Ryabtsev

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["ryabtsev"]
ind2["ryabtsev"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "ryabtsev")$value
dp(s = s2, v = v2, measure = "ryabtsev")$value
```
:::
::::

### Szalai

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["szalai"]
ind2["szalai"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "szalai")$value
dp(s = s2, v = v2, measure = "szalai")$value
```
:::
::::

### Weighted Szalai

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["w_szalai"]
ind2["w_szalai"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "weighted szalai")$value
dp(s = s2, v = v2, measure = "weighted szalai")$value
```
:::
::::

### Aleskerov & Platonov

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["ap"]
ind2["ap"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "aleskerov")$value
dp(s = s2, v = v2, measure = "aleskerov")$value
```
:::
::::

### Gini coefficient

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["gini"]
ind2["gini"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "gini")$value
dp(s = s2, v = v2, measure = "gini")$value
```
:::
::::

### Atkinson

* Parameter $\eta$ (`eta`) is required. Defualt is 2.

::::{.columns}
:::{.column width=49%}
```{r}
#| filename: "{PRcalc}"
index(obj1, eta = 2)["atkinson"]
index(obj2, eta = 3)["atkinson"]
```
:::

:::{.column width=2%}
:::

:::{.column width=49%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, eta = 2, 
   measure = "atkinson")$value
dp(s = s2, v = v2, eta = 3, 
   measure = "atkinson")$value
```
:::
::::

### Generalized entropy

See $\alpha$-divergence.

### Sainte-Laguë

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["sl"]
ind2["sl"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "sainte lague")$value
dp(s = s2, v = v2, measure = "sainte lague")$value
```
:::
::::

### Cox & Shugart

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["cs"]
ind2["cs"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "cox shugart")$value
dp(s = s2, v = v2, measure = "cox shugart")$value
```
:::
::::

### Farina

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["farina"]
ind2["farina"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "farina")$value
dp(s = s2, v = v2, measure = "farina")$value
```
:::
::::

### Ortona

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["ortona"]
ind2["ortona"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "ortona")$value
dp(s = s2, v = v2, measure = "ortona")$value
```
:::
::::

### Cosine Dissimilarity

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["cd"]
ind2["cd"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "cosine")$value
dp(s = s2, v = v2, measure = "cosine")$value
```
:::
::::

### Lebeda's RR (Mixture D'Hondt)

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["rr"]
ind2["rr"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "mixture")$value
dp(s = s2, v = v2, measure = "mixture")$value
```
:::
::::

### Lebeda's ARR

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["arr"]
ind2["arr"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "arr")$value
dp(s = s2, v = v2, measure = "arr")$value
```
:::
::::

### Lebeda's SRR

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["srr"]
ind2["srr"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "srr")$value
dp(s = s2, v = v2, measure = "srr")$value
```
:::
::::

### Lebeda's WDRR

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["wdrr"]
ind2["wdrr"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "wdrr")$value
dp(s = s2, v = v2, measure = "wdrr")$value
```
:::
::::

### Kullback-Leibler Surprise

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["kl"]
ind2["kl"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "surprise")$value
dp(s = s2, v = v2, measure = "surprise")$value
```
:::
::::

### Likelihood Ratio Statistic

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["lr"]
ind2["lr"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "lrstat")$value
dp(s = s2, v = v2, measure = "lrstat")$value
```
:::
::::

### $\chi^2$

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["chisq"]
ind2["chisq"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "chisq")$value
dp(s = s2, v = v2, measure = "chisq")$value
```
:::
::::

### Hellinger Distance

::::{.columns}
:::{.column width=29%}
```{r}
#| filename: "{PRcalc}"
ind1["hellinger"]
ind2["hellinger"]
```
:::

:::{.column width=2%}
:::

:::{.column width=69%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, measure = "hellinger")$value
dp(s = s2, v = v2, measure = "hellinger")$value
```
:::
::::

### $\alpha$-divergence (Generalized entropy)

* Parameter $\alpha$ (`alpha`) is required. Default is 2.
* $\alpha$-divergence is theoretically equivalent to generalized entropy. However, {seatdist} can only be computed for the cases where $\alpha \neq 0, 1$

::::{.columns}
:::{.column width=49%}
```{r}
#| filename: "{PRcalc}"
index(obj1, alpha = 2)["ad"]
index(obj2, alpha = 0, as_disprop = FALSE)["ad"]
index(obj1, alpha = 1)["ad"]
index(obj2, alpha = -1, as_disprop = FALSE)["ad"]
index(obj1, alpha = 3)["ad"]
index(obj2, alpha = 0.5)["ad"]
```
:::

:::{.column width=2%}
:::

:::{.column width=49%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, alpha = 2, 
   measure = "gen entropy")$value
dp(s = s2, v = v2, alpha = 0, 
   measure = "gen entropy")$value
dp(s = s1, v = v1, alpha = 1, 
   measure = "gen entropy")$value
dp(s = s2, v = v2, alpha = -1, 
   measure = "gen entropy")$value
dp(s = s1, v = v1, alpha = 3, 
   measure = "gen entropy")$value
dp(s = s2, v = v2, alpha = 0.5, 
   measure = "gen entropy")$value
```
:::
::::

**Example 1:** $D^{-1} \times 2$ is equivalent to Neyman's $\chi^2$ divergence.

```{r}
2 * index(obj2, alpha = -1, as_disprop = FALSE)["ad"]
index(obj2)["chisq"]
```

**Example 2:** $D^{1}$ is equivalent to Kullback-Leibler surprise.

```{r}
index(obj2, alpha = 1, as_disprop = FALSE)["ad"]
index(obj2)["kl"]
```
