---
title: "Decomposition"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
toc: true
---

## Preparation

```{r}
#| label: setup
#| message: false
library(PRcalc)
library(tidyverse)
```

## Two-step decomposition

$\alpha$-divergence ($D^\alpha$) can be decomposed in two components---apportionment and districting as follow:

$$
\begin{align}
D^{\alpha} = & \sum_{j = 1}^{k} v_j \frac{1}{\alpha (\alpha - 1)} \log{\frac{s_j}{v_j}} + \\
& \sum_{j=1}^{k} (s_j)^\alpha (v_j)^{1 - \alpha} \sum_{i=1}^{k_j} v_{ji} \frac{1}{\alpha (\alpha - 1)} \Biggl[\biggl(\frac{s_{ji}}{v_{ji}}\biggr)^{\alpha} - 1 \Biggr]\\\
= & D^{\alpha}_{\mbox{apportionment}} + \sum_{j=1}^k W^{\alpha}_j D^{\alpha}_{\mbox{districting}_j}.
\end{align}
$$

However there are two special cases, $\alpha \rightarrow 0$ and $\alpha \rightarrow 1$.

$$
\begin{align}
D^0 = & \sum_{j = 1}^{k} v_j \log{\frac{v_j}{s_j}} + \sum_{j=1}^{k} v_j \sum_{i=1}^{k_j} v_{ji} \log{\frac{v_{ji}}{s_{ji}}} \\
= & D^0_{\mbox{apportionment}} + \sum_{j=1}^k W^0_j D^0_{\mbox{districting}_j}.
\end{align}
$$

$$
\begin{align}
D^1 = & \sum_{j = 1}^{k} s_j \log{\frac{s_j}{v_j}} + \sum_{j=1}^{k} s_j \sum_{i=1}^{k_j} s_{ji} \log{\frac{s_{ji}}{v_{ji}}} \\
= & D^1_{\mbox{apportionment}} + \sum_{j=1}^k W^1_j D^1_{\mbox{districting}_j}.
\end{align}
$$

### Using a `prcalc` object.

```{r}
data("jp_lower_2021_en")

obj1 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 22, 17, 11, 21, 28, 11, 6, 20),
               method = "dt")

obj2 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 23, 19, 11, 20, 29, 10, 5, 19),
               method = "dt")

obj3 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 23, 19, 11, 20, 29, 10, 5, 19),
               method = "msl")

decompose(obj1)
decompose(obj2)
decompose(obj3)
```

Different $\alpha$ also available. Default is 2.

```{r}
decompose(obj1, alpha = 0.5)
decompose(obj1, alpha = 1)
decompose(obj1, alpha = 1.5)
decompose(obj1, alpha = 2)
decompose(obj1, alpha = 3)
```

### Import `prcalc` object and decomposition.

#### From `.csv` format

Sample dataset can be downloaded from [here](../data/sample_dataset.csv).

```{r}
obj4 <- read_prcalc("../data/sample_dataset.csv",
                    l1   = "region",
                    l2   = "district",
                    p    = "electorates",
                    q    = "magnitude",
                    type = "district")

print(obj4, head = 10)
decompose(obj4, alpha = 2)
decompose(obj4, alpha = 0)
```


#### From `data.frame`

```{r}
data(au_district_2010)

head(au_district_2010)

# If all magnitudes are 1, "q" can be omitted.
obj5 <- as_prcalc(au_district_2010,
                  l1   = "region",
                  l2   = "district",
                  p    = "electorates",
                  type = "district")

print(obj5, head = 10)

decompose(obj5, alpha = 2)
decompose(obj5, alpha = 0)
```

```{r}
data(jp_lower_2021_result)

head(jp_lower_2021_result)

obj6 <- as_prcalc(jp_lower_2021_result,
                  l1   = "Pref",
                  l2   = "Party",
                  p    = "Votes",
                  q    = "Seats",
                  type = "party")

print(obj6, head = 10)

decompose(obj6, alpha = 2)
decompose(obj6, alpha = 0)
```

## Three-step Decomposition

If `special` is defined, three-step decomposition is conducted.

```{r}
data("nz_district")

nz_1949 <- nz_district |>
  dplyr::filter(year == 1949) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "district")

nz_1972 <- nz_district |>
  dplyr::filter(year == 1972) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "district")

nz_1990 <- nz_district |>
  dplyr::filter(year == 1990) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "district")

nz_2011 <- nz_district |>
  dplyr::filter(year == 2011) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "district")

(decompose_1949 <- decompose(nz_1949, alpha = 1, special = "Maori"))
(decompose_1972 <- decompose(nz_1972, alpha = 1, special = "Maori"))
(decompose_1990 <- decompose(nz_1990, alpha = 1, special = "Maori"))
(decompose_2011 <- decompose(nz_2011, alpha = 1, special = "Maori"))

list("1949" = decompose_1949,
     "1972" = decompose_1972,
     "1990" = decompose_1990,
     "2011" = decompose_2011) |> 
  compare() 

list("1949" = decompose_1949,
     "1972" = decompose_1972,
     "1990" = decompose_1990,
     "2011" = decompose_2011) |> 
  compare() |> 
  plot() +
  labs(title = "Electoral districts in New Zealand",
       caption = expression(paste("Three-step decomposition with ", alpha, " = 1")))
```

```{r}
# 1949-2011
nz_decompose_all <- nz_district |> 
  group_by(year) |> 
  nest() |> 
  mutate(data      = map(data, \(x) as_prcalc(x,
                                              l1   = "region",
                                              l2   = "district",
                                              p    = "electorates",
                                              q    = "magnitude",
                                              type = "district")),
         decompose = map(data, \(x) decompose(x, 
                                              alpha      = 0,
                                              special    = "Maori"))) |> 
  pull(decompose)

names(nz_decompose_all) <- unique(nz_district$year)

nz_decompose_all |> 
  compare() |> 
  print(use_gt = TRUE, digits = 5)
```

```{r}
#| fig-width: 8
#| fig-height: 6
nz_decompose_all |> 
  compare() |> 
  plot(value_type = "total", value_angle = 90, 
       x_angle = 45, border_width = 0.5, digits = 6, border_color = "white") +
  coord_cartesian(ylim = c(0, 0.025)) +
  labs(title = "Electoral districts in New Zealand (1949-2011)",
       caption = expression(paste("Note: Three-step decomposition with ", alpha, " = 2")))
```

```{r}
#| fig-width: 6
#| fig-height: 18
nz_decompose_all |> 
  compare() |> 
  plot(facet = TRUE, value_size = 2, digits = 5)
```

## Comparison

```{r}
#| label: comparison-decomposition
compare(list("D'Hondt"                           = decompose(obj1),
             "D'Hondt + Modified magnitudes"     = decompose(obj2),
             "Modified-SL + Modified magnitudes" = decompose(obj3))) 

compare(list("D'Hondt"                          = decompose(obj1),
             "D'Hondt\nModified magnitudes"     = decompose(obj2),
             "Modified-SL\nModified magnitudes" = decompose(obj3)))  |> 
  plot()

compare(list("D'Hondt"                           = decompose(obj1),
             "D'Hondt + Modified magnitudes"     = decompose(obj2),
             "Modified-SL + Modified magnitudes" = decompose(obj3))) |>
  plot(facet = TRUE, value_type = "text", digits = 5)
```

## Comparison between `index()` and `decompose()`

Please see "Marginal and joint distribution" in [*Measuring disproportionality*](disproportionality.qmd).
