---
title: "Decomposition"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
toc: true
---

## Preparation

```{r}
#| label: setup
#| message: false
library(PRcalc)
library(tidyverse)
```

## Two-step decomposition

$\alpha$-divergence ($D^\alpha$) can be decomposed in two components---apportionment and districting as follow:

$$
\begin{align}
D^{\alpha} = & \underbrace{\sum_{j = 1}^{k} v_j \frac{1}{\alpha (\alpha - 1)} \log{\frac{s_j}{v_j}}}_{\mbox{approtionment term}} + \\
& \underbrace{\sum_{j=1}^{k} (s_j)^\alpha (v_j)^{1 - \alpha} \sum_{i=1}^{k_j} v_{j_i} \frac{1}{\alpha (\alpha - 1)} \Biggl[\biggl(\frac{s_{j_i}}{v_{j_i}}\biggr)^{\alpha} - 1 \Biggr]}_{\mbox{districting term}}\\\
= & D^{\alpha}_{\mbox{apportionment}} + \sum_{j=1}^k W^{\alpha}_j D^{\alpha}_{\mbox{districting}_j}.
\end{align}
$$

However there are two special cases, $\alpha \rightarrow 0$ and $\alpha \rightarrow 1$.

$$
\begin{align}
D^0 = & \underbrace{\sum_{j = 1}^{k} v_j \log{\frac{v_j}{s_j}}}_{\mbox{approtionment}} + \underbrace{\sum_{j=1}^{k} v_j \sum_{i=1}^{k_j} v_{j_i} \log{\frac{v_{j_i}}{s_{j_i}}}}_{\mbox{districting}} \\
= & D^0_{\mbox{apportionment}} + \sum_{j=1}^k W^0_j D^0_{\mbox{districting}_j}.
\end{align}
$$

$$
\begin{align}
D^1 = & \underbrace{\sum_{j = 1}^{k} s_j \log{\frac{s_j}{v_j}}}_{\mbox{approtionment}} + \underbrace{\sum_{j=1}^{k} s_j \sum_{i=1}^{k_j} s_{j_i} \log{\frac{s_{j_i}}{v_{j_i}}}}_{\mbox{districting}} \\
= & D^1_{\mbox{apportionment}} + \sum_{j=1}^k W^1_j D^1_{\mbox{districting}_j}.
\end{align}
$$

```{r}
data("jp_lower_2021_en")

obj1 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 22, 17, 11, 21, 28, 11, 6, 20),
               method = "dt")

obj2 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 23, 19, 11, 20, 29, 10, 5, 19),
               method = "dt")

obj3 <- prcalc(jp_lower_2021_en,
               m = c(8, 13, 19, 23, 19, 11, 20, 29, 10, 5, 19),
               method = "msl")

decompose(obj1)
decompose(obj2)
decompose(obj3)
```

Different $\alpha$ also available. Default is 2.

:::{.panel-tabset}
## $\alpha$ = 0.5

```{r}
decompose(obj1, alpha = 0.5)
```

## $\alpha$ = 1

```{r}
decompose(obj1, alpha = 1)
```

## $\alpha$ = 2

```{r}
decompose(obj1, alpha = 2)
```

## $\alpha$ = 3

```{r}
decompose(obj1, alpha = 3)
```
:::

## Three-step Decomposition

If there is a level 0, such as overseas constituencies or minority constituencies, $\alpha$-divergence ($D^\alpha$) can be decomposed into three components---special, apportionment and districting terms as follow:

$$
\begin{align}
D^{\alpha} = & 
\underbrace{\sum_{h = 1}^{k} v_h \frac{1}{\alpha (\alpha - 1)} \Biggl[\biggl(\frac{s_{h}}{v_{h}}\biggr)^{\alpha} - 1 \Biggr]}_{\mbox{special}} + \\
& \underbrace{\sum_{h = 1}^{k} (s_h)^{\alpha} (v_h)^{1 - \alpha} \sum_{j=1}^{k_j} v_{h_j} \frac{1}{\alpha (\alpha - 1)} \Biggl[\biggl(\frac{s_{h_j}}{v_{h_j}}\biggr)^{\alpha} - 1 \Biggr]}_{\mbox{approtionment}} \\
& \underbrace{\sum_{h = 1}^{k} (s_h)^{\alpha} (v_h)^{1 - \alpha} \sum_{j=1}^{k_j} (s_{h_j})^{\alpha} (v_{h_j})^{1 - \alpha} \sum_{i=1}^{k_{h_j}} v_{h_{j_i}} \frac{1}{\alpha (\alpha - 1)} \Biggl[\biggl(\frac{s_{h_{j_i}}}{v_{h_{j_i}}}\biggr)^{\alpha} - 1 \Biggr]}_{\mbox{districting}} \\
= & D^{\alpha}_{\mbox{special}} + \sum_{h=1}^{k}W_h^{\alpha} D^{\alpha}_{\mbox{apportionment}_h} + \sum_{h=1}^{k} \sum_{j=1}^{k_h} w_{h_j}^{\alpha} D^{\alpha}_{\mbox{districting}_{h_j}}.
\end{align}
$$

However there are two special cases, $\alpha \rightarrow 0$ and $\alpha \rightarrow 1$.

$$
\begin{align}
D^0 = & 
\underbrace{\sum_{h=1}^{k} v_h \biggl(\log{\frac{v_h}{s_h}}\biggr)}_{\mbox{special}} + 
\underbrace{\sum_{h=1}^{k} v_h \sum_{j=1}^{k_h} v_{h_j} \biggl(\log{\frac{v_{h_j}}{s_{h_j}}}\biggr)}_{\mbox{apportionment}} + \\
& \underbrace{\sum_{h=1}^{k} v_h \sum_{j=1}^{k_h} v_{h_j} \sum_{i=1}^{k_{h_j}} \biggl(\log{\frac{v_{h_{j_i}}}{s_{h_{j_i}}}}\biggr)}_{\mbox{districting}} \\
= & D^0_{\mbox{special}} + \sum_{h=1}^{k}W_h^0 D^0_{\mbox{apportionment}_h} + \sum_{h=1}^{k} \sum_{j=1}^{k_h} w_{h_j}^0 D^0_{\mbox{districting}_{h_j}}.
\end{align}
$$

$$
\begin{align}
D^1 = & 
\underbrace{\sum_{h=1}^{k} s_h \biggl(\log{\frac{s_h}{v_h}}\biggr)}_{\mbox{special}} + 
\underbrace{\sum_{h=1}^{k} s_h \sum_{j=1}^{k_h} v_{h_j} \biggl(\log{\frac{s_{h_j}}{v_{h_j}}}\biggr)}_{\mbox{apportionment}} + \\
& \underbrace{\sum_{h=1}^{k} s_h \sum_{j=1}^{k_h} s_{h_j} \sum_{i=1}^{k_{h_j}} \biggl(\log{\frac{s_{h_{j_i}}}{v_{h_{j_i}}}}\biggr)}_{\mbox{districting}} \\
= & D^1_{\mbox{special}} + \sum_{h=1}^{k}W_h^1 D^1_{\mbox{apportionment}_h} + \sum_{h=1}^{k} \sum_{j=1}^{k_h} w_{h_j}^1 D^1_{\mbox{districting}_{h_j}}.
\end{align}
$$

A `special` argument must be specified for three-step decomposition. In the case of New Zealand in 1949, four Maori electoral districts were in the special region, `"Maori"`.

```{r}
data("nz_district")

nz_1949 <- nz_district |>
  dplyr::filter(year == 1949) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "nested")

nz_1949
```

These four Maori electoral districts are treated as a special region (`"Maori"`) and assigned to `special` argument (`special = "Maori"`).

```{r}
decompose(nz_1949, alpha = 1, special = "Maori")
```

## Comparison

```{r}
#| label: comparison-decomposition
compare(list("D'Hondt"                           = decompose(obj1),
             "D'Hondt + Modified magnitudes"     = decompose(obj2),
             "Modified-SL + Modified magnitudes" = decompose(obj3))) 

list("Scenario 1" = decompose(obj1),
     "Scenario 2" = decompose(obj2),
     "Scenario 3" = decompose(obj3)) |> 
  compare() |> 
  print(use_gt = TRUE, digits = 5)
```

## Visualization

後ほど、{purrr}を使ったより効率的なやり方を紹介する。

```{r}
nz_1972 <- nz_district |>
  dplyr::filter(year == 1972) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "nested")

nz_1990 <- nz_district |>
  dplyr::filter(year == 1990) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "nested")

nz_2011 <- nz_district |>
  dplyr::filter(year == 2011) |>
  as_prcalc(l1   = "region",
            l2   = "district",
            p    = "electorates",
            q    = "magnitude",
            type = "nested")

decompose_1949 <- decompose(nz_1949, alpha = 1, special = "Maori")
decompose_1972 <- decompose(nz_1972, alpha = 1, special = "Maori")
decompose_1990 <- decompose(nz_1990, alpha = 1, special = "Maori")
decompose_2011 <- decompose(nz_2011, alpha = 1, special = "Maori")

nz_compare <- list("1949" = decompose_1949,
                   "1972" = decompose_1972,
                   "1990" = decompose_1990,
                   "2011" = decompose_2011) |> 
  compare()

print(nz_compare, use_gt = TRUE, digits = 5)
```

```{r}
nz_compare |> 
  plot() +
  labs(x = "Year (D1)",
       title = "Electoral districts in New Zealand",
       caption = expression(paste("Three-step decomposition with ", alpha, " = 1")))

list("1949" = decompose_1949,
     "1972" = decompose_1972,
     "1990" = decompose_1990,
     "2011" = decompose_2011) |> 
  compare() |> 
  plot(percentage = TRUE) +
  labs(x = "Year (D1)", y = "(%)",
       title = "Electoral districts in New Zealand",
       caption = expression(paste("Three-step decomposition with ", alpha, " = 1")))
```

## Compatibility with {purrr}

```{r}
# 1949-2011
nz_decompose_all <- nz_district |> 
  group_by(year) |> 
  nest() |> 
  mutate(data      = map(data, \(x) as_prcalc(x,
                                              l1   = "region",
                                              l2   = "district",
                                              p    = "electorates",
                                              q    = "magnitude",
                                              type = "nested")),
         decompose = map(data, \(x) decompose(x, 
                                              alpha      = 0,
                                              special    = "Maori"))) |> 
  pull(decompose)

names(nz_decompose_all) <- unique(nz_district$year)

nz_decompose_all |> 
  compare() |> 
  print(use_gt = TRUE, digits = 5)
```

```{r}
#| fig-width: 8
#| fig-height: 6
nz_decompose_all |> 
  compare() |> 
  plot(value_type = "total", value_angle = 90, 
       x_angle = 45, border_width = 0.5, digits = 6, border_color = "white") +
  coord_cartesian(ylim = c(0, 0.025)) +
  labs(title = "Electoral districts in New Zealand (1949-2011)",
       caption = expression(paste("Note: Three-step decomposition with ", alpha, " = 2")))
```

```{r}
#| fig-width: 6
#| fig-height: 18
nz_decompose_all |> 
  compare() |> 
  plot(facet = TRUE, value_size = 2, digits = 5)
```

## Comparison between `index()` and `decompose()`

Please see "Marginal and joint distribution" in [*Measuring disproportionality*](disproportionality.qmd).
