---
title: "List of methods"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
---

## Highest averages methods

* $r(k)$: $k+1$th divisor

|Method|Argument|Rule|Approx. first values|
|:---|:---|:---|:---|
|D'hondt / Jefferson |`"d'hondt"`, `"dt"`, or `"jefferson"`|$r(k) = k + 1$|1, 2, 3, 4, ...|
|Adams / Cambridge|`"adams"` or `"cambridge"`|$r(k) = k$|0, 1, 2, 3, ...|
|Sainte-Laguë / Webster|`"sl"`, `"sainte-lague"`, or `"webster"`|$r(k) = k + 0.5$|0.5, 1.5, 2.5, 3.5, ...|
|Modified Sainte-Laguë|`"msl"`|$r(k) = k + 0.5, r_1 = 0.7$|0.7, 1.5, 2.5, 3.5, ...|
|Danish|`"danish"`|$r(k) = k + \frac{1}{3}$|0.33, 1.33, 2.33, 3.33, ...|
|Imperiali|`"imperiali"`|$r(k) = k + 2$|2, 3, 4, 5, ...|
|Huntington-Hill|`"hh"` or `"huntington-hill"`|$r(k) = \sqrt{k (k + 1)}$|0, 1.41, 2.45, 3.46, ...|
|Dean|`"dean"`|$r(k) = \frac{2}{\frac{1}{k} + \frac{1}{k + 1}}$|0, 1.33, 2.40, 3.43, ...|
|$\alpha$-divergence|`"ad"`|See @nte-ad |See @nte-ad |
|Custom divisor|`"custom divisor"` or `"cd"`|See @nte-cd |See @nte-cd |

## Largest reminder method

* $Q$: Quota
* $v$: Total votes
* $m$: Magnitude (total seats)

|Method|Argument|Quota|
|:---|:---|:---|:---|
|Hare-Niemeyer quota|`"hn"`, `"hare"`, or `"hare-niemeyer"`|$Q = \frac{v}{m}$|
|Droop quota|`"droop"`|$Q = 1 + \frac{v}{1 + m}$|
|Hagenbach-Bischoff quota|`"hb"` or `"hagenbach-bischoff"`|$Q = \frac{v}{1 + m}$|
|Imperiali quota|`"iq"` or `"imperiali quota"`|$Q = \frac{v}{2 + m}$|
|Custom quota|`"custom quota"` or `"cq"`|See @nte-cq |

:::{#nte-ad .callout-note collapse="true"}
## About $\alpha$-divergence

Divisor ($r(k)$) is calculated as follow:

$$
r(k) = \begin{cases}
\frac{1}{e} \frac{(k + 1)^{k + 1}}{k^k}& \mbox{if}\quad\alpha = 1,\\
\frac{1}{\log{\frac{k + 1}{k}}} & \mbox{if}\quad\alpha = 0,\\
\Bigl(\frac{(k + 1)^\alpha - k^\alpha}{\alpha}\Bigr)^{\frac{1}{\alpha - 1}} & \mbox{otherwise.}\\
\end{cases}
$$

|$\alpha$|Method|
|---:|:---|
|+$\infty$|D'hondt / Jefferson|
|3|$\frac{1}{3}$|
|2|Sainte-Laguë / Webster|
|1|Theil|
|0|Theil and Schrage|
|-1|Hill|
|-4|$\simeq$Dean|
|-$\infty$|Adams / Cambridge|

Please note that extra parameter $\alpha$ is required (`extra`). 

:::{.panel-tabset}
## Example 1

```{r}
#| message: false
library(PRcalc)
data(jp_upper_2019)
# Jefferson vs alpha = 10
list(prcalc(jp_upper_2019, m = 50, method = "dt"),
     prcalc(jp_upper_2019, m = 50, method = "ad", extra = 10)) |> 
  compare() |> 
  print(use_gt = TRUE)
```

## Example 2

```{r}
#| message: false
# Dean vs alpha = -4
list(prcalc(jp_upper_2019, m = 50, method = "dean"),
     prcalc(jp_upper_2019, m = 50, method = "ad", extra = -4)) |> 
  compare() |> 
  print(use_gt = TRUE)
```
:::

:::

:::{#nte-cd .callout-note collapse="true"}
## About `"custom divisor"` (`"cd"`) method

Please note that extra parameter (`extra`) is required. `extra` must be length of maximum value of `m` or longer.

```{r}
#| message: false
list(prcalc(jp_upper_2019, m = 50, method = "msl"),
     prcalc(jp_upper_2019, m = 50, method = "cd", 
            extra = c(1.4, seq(3, 100, by = 2)))) |> 
  compare() |> 
  print(use_gt = TRUE)
```

For detail, see "[Customized allocation methods](../tutorials/calculation.html#customized-allocation-methods)" section in calculation tutorial.
:::

:::{#nte-cq .callout-note collapse="true"}
## About `"custom quota"` (`"cq"`) method

Please note that extra parameter (`extra`) is required. `extra` must be length of maximum value of `m` or longer. For detail, see "[Customized allocation methods](../tutorials/calculation.html#customized-allocation-methods)" section in calculation tutorial.

:::
