---
title: "Comparison with other packages"
knitr: 
  opts_chunk: 
    dev: "ragg_png"
    fig.width: 7
    fig.height: 5
    dpi: 150
    collapse: true
---

## Setting

```{r}
#| message: false
library(tidyverse)
library(PRcalc)
library(seatdist)

dp <- disproportionality

obj1 <- prcalc(jp_upper_2019, m = 50, method = "dt")
v1   <- obj1$raw$Vote
s1   <- obj1$dist$Vote

obj2 <- prcalc(us_census_2020, m = 435, method = "hh")
v2   <- obj2$raw$Pop
s2   <- obj2$dist$Pop
```

## Comparison

### D'Hondt

::::{.columns}
:::{.column width=49%}
```{r}
#| filename: "{PRcalc}"
index(obj1)["dhondt"]
index(obj2)["dhondt"]
```
:::

:::{.column width=2%}
:::

:::{.column width=49%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, 
   measure = "dhondt")$value
dp(s = s2, v = v2, 
   measure = "dhondt")$value
```
:::
::::

### Monroe

::::{.columns}
:::{.column width=49%}
```{r}
#| filename: "{PRcalc}"
index(obj1)["monroe"]
index(obj2)["monroe"]
```
:::

:::{.column width=2%}
:::

:::{.column width=49%}
```{r}
#| filename: "{seatdist}"
dp(s = s1, v = v1, 
   measure = "monroe")$value
dp(s = s2, v = v2, 
   measure = "monroe")$value
```
:::
::::

### Maximum absolute deviation

### Rae

### Loosemore & Hanby

### Grofman

### Lijphart

### Gallagher

### Generalized Gallagher

* $k$

### Gatev

### Ryabtsev

### Szalai

### Weighted Szalai

### Aleskerov & Platonov

### Gini coefficient

### Atkinson

* $\eta$

### Sainte-Laguë

### Cox & Shugart

### Farina

### Ortona

### Cosine Dissimilarity

### Lebeda's RR (Mixture D'Hondt)

### Lebeda's ARR

### Lebeda's SRR

### Lebeda's WDRR

### Kullback-Leibler Surprise

### Likelihood Ratio Statistic

### $\chi^2$

### Hellinger Distance

### $\alpha$-divergence (Generalized entropy)

* $\alpha$
* {seatdist} can only be computed for the cases where $\alpha \neq 0, 1$
